##Loading and preprocessing the data
1. Load the data
```{r}
library(dplyr)
library(tidyr) #may need it later

file = read.csv("D:/Cours/Coursera/Reproducible research/Week2/activity.csv")
```
2. Transform the data into a format suitable for your analysis
```{r}
data = tbl_df(file)
```


##What is mean total number of steps taken per day?
Calculate the total number of steps taken per day
```{r}
library(ggplot2)
steps <- data %>% 
    na.omit() %>% #removing the NAs
    group_by(date) %>%
    summarise("number_of_steps" = sum(steps))
```

2. Plotting the histogram
```{r}
g <- ggplot(steps, aes(date)) + 
    geom_histogram(bins=53, aes(number_of_steps)) + 
    theme_bw() +
    ggtitle("Histogram of the frenquence of the number of steps") + 
    theme(plot.title = element_text(hjust = 0.5)) +
    xlab("Number of steps") + 
    ylab("frequency")

print(g)
```

3.Compute and reporting the mean and the median
```{r}
meanSteps = mean(steps$number_of_steps)
medianSteps = median(steps$number_of_steps)
g  +
    geom_vline(xintercept = meanSteps, col="red", show.legend = TRUE) +
    geom_vline(xintercept = medianSteps, col = "green", show.legend = TRUE) 
    
```

* Mean = `r meanSteps`
* Median = `r medianSteps`
In my data, the mean and the median are so close that they are not distinct on the graph, for some reason won't display the labels on the graph

##What is the average daily activity pattern?
```{r}
stepsbyinterval <- data %>%
    na.omit() %>%
    group_by(interval) %>%
    summarise("number_of_steps" = mean(steps))
```

1. Time series plot
```{r}
ggplot(stepsbyinterval, aes(x=interval, y=number_of_steps)) +
    geom_line() +
    theme_bw() + 
    ggtitle("Time series of the number of the mean(steps)") + 
    theme(plot.title = element_text(hjust = 0.5)) +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxSteps <- stepsbyinterval[which.max(stepsbyinterval$number_of_steps), 1]
```
The interval number **`r maxSteps`** contains the maximum number of steps


##Imputing missing values
1.Calculate and report the total number of missing values in the dataset
```{r}
sumNAs = sum(is.na(data))
```
There are `r sumNAs` missing values in the dataset.

2. and 3. Filling in all of the missing values in the dataset
```{r}
#filling NAs with the mean of the interval
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
filledData <- data %>%
    group_by(interval) %>%
    mutate(steps=impute.mean(steps)) %>%
    ungroup
```
The strategy used to fill the missing values is using the mean for each interval


4.
```{r}

stepsPerDay <- filledData %>%
    group_by(date) %>%
    summarise("number_of_steps" = sum(steps))
    
meanStepsFilled <- mean(stepsPerDay$number_of_steps)
medianStepsFilled <- median(stepsPerDay$number_of_steps)
    
```
* Mean = `r meanStepsFilled`
* Median = `r medianStepsFilled`

Since we imputed the missing values with the means / interval, the overall mean is still the same.

```{r}
ggplot(stepsPerDay, aes(date)) + 
    geom_histogram(bins=61, aes(number_of_steps)) + 
    theme_bw() +
    ggtitle("Histogram of the frenquence of the number of steps with NA imputed") + 
    theme(plot.title = element_text(hjust = 0.5)) +
    xlab("Number of steps") + 
    ylab("frequency") +
    geom_vline(xintercept = meanStepsFilled, col="red", show.legend = TRUE) +
    geom_vline(xintercept = medianStepsFilled, col = "green", show.legend = TRUE) 
```


##Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend”
```{r}
#indices of days of the weekend
library(lubridate)
filledData2 <- filledData

filledData2$date <- as.Date(filledData2$date) 
filledData2$day = ifelse(wday(filledData2$date-1) <6, "weekday", "weekend")
filledData2$day = as.factor(filledData2$day)

```
2. Time series plot
```{r}

filledData2 <- filledData2 %>%
    group_by(day, interval) %>%
    summarize(meanSteps = mean(steps))

ggplot(filledData2, aes(x=interval, y=meanSteps)) + geom_line() + facet_grid(day ~.) + ylab("Average number of steps") +
    ggtitle("Daily activity for week days and weekends") + 
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_bw()
```

As we can see, there is much less activity during the weekend.